---
title: "demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nsphsmlqt)
```

# Input

We use whole genome sequence. Here is a sample of genome sequences:

```{r genetic_data}
knitr::kable(head(get_genome_sequences()))
```

Here is a sample of the phenotypic data:

```{r phenotypic_data}
knitr::kable(head(get_traits()))
```

# Select relevant SNVs

Extract the SNVs from genetic data:

```{r snvs}
print("First 10 SNVs")
```

Annotate the SNVs:

```{r snvs_with_annotation}
print("First 10 SNVs, with gene name and location")
```

We now have [an amount] of SNV.

List of genes that are associated with CADs:

```{r cad_genes}
print("t_cad <- knitr::kabel(get_cad_associated_gene_names())")
```

Select SNVs that have genes association with CADs

```{r snvs_related_to_cads}
print("select SNV with name in t_cad")
```

We now have [a lower amount] of SNV.


# Use BEDMatrix to read .bed
