---
title: "snps_used"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{snps_used}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Use on of these two:
# output: rmarkdown::html_vignette
# output: rmarkdown::pdf_document

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

> This document is a vignette of the R package `nsphs_ml_qt`
> and can be found at [https://github.com/richelbilderbeek/nsphs_ml_qt/blob/master/vignettes/snps_used.Rmd](https://github.com/richelbilderbeek/nsphs_ml_qt/blob/master/vignettes/snps_used.Rmd).

```{r setup}
library(nsphsmlqt)
```

In this study, in the exploratory phase, we used two SNPs.

First, we get Table 2 from Kierczak et al., 2022, and show the top 
rows (this will look messy!):

```{r}
t <- get_kierczak_et_al_2022_table_s2_xlsx()
knitr::kable(head(t))
```

We will use the default columns names from this table, even
though these names do not follow a Tidyverse naming style.

To fix the messiness due to having too much columnsm, 
we'll only use the columns we need:

```{r}
t <- dplyr::select(
  t, 
  X.Primary.Conditional.hit, 
  Protein, 
  Panel_Position,
  Cis_Trans,
  Chr,
  SNP,
  Position., # nolint period at the end of 'Position', as in original data
  P.value
)
# p-values are too small to display in default notation
t$P.value <- format(t$P.value, scientific = TRUE)
knitr::kable(head(t))
```


## Finding the SNP with the strongest assocation

Finding the lowest p-value results in protein IL-6RA:

```{r}
lowest_p_value <- min(t$P.value)
row_with_lowest_p_value <- which(t$P.value == lowest_p_value)
t_with_lowest_p_value <- t[row_with_lowest_p_value, ]
testthat::expect_equal(t_with_lowest_p_value$Protein, "IL-6RA")
testthat::expect_equal(t_with_lowest_p_value$Panel_Position, "CVD3_142")
knitr::kable(t_with_lowest_p_value)
```

IL-6RA has three signals in total:

```{r}
t_is_6ra <- t[which(t$Protein == "IL-6RA"), ]
testthat::expect_equal(nrow(t_is_6ra), 3)
knitr::kable(t_is_6ra)
```

## Finding the SNP with the most cis assocations

Find the number of assocations per protein:

```{r}
t_n_signals <- dplyr::summarise(
  dplyr::group_by(
    dplyr::select(t, "Protein", "Cis_Trans"), 
    Protein
  ), 
  n = dplyr::n()
)
knitr::kable(head(t_n_signals))
```

Find the protein with the most associations:

```{r}
t_most_signals <- t_n_signals[which(t_n_signals$n == max(t_n_signals$n)), ]
testthat::expect_equal(t_most_signals$Protein, "IL-17RA")
knitr::kable(t_most_signals)
```

So IL-17RA has the most associations, let's see all:

```{r}
t_il_17ra <- t[t$Protein == t_most_signals$Protein, ]
testthat::expect_equal(nrow(t_il_17ra), 9)
knitr::kable(t_il_17ra)
```

So IL-17RA has 9 associations. 
Here is how these are distributed over the chromosomes:

```{r}
t_n_signals_per_chromosome <- dplyr::summarise(
  dplyr::group_by(
    dplyr::select(t_il_17ra, "Chr"), 
    Chr
  ), 
  n = dplyr::n()
)
knitr::kable(head(t_n_signals_per_chromosome))
```

Here we pick the chromosome with the most signals:

```{r}
row_with_max_signals <- which(
  t_n_signals_per_chromosome$n == max(t_n_signals_per_chromosome$n)
)
chromosome_with_most_signals <- t_n_signals_per_chromosome[row_with_max_signals, ]$Chr
testthat::expect_equal(chromosome_with_most_signals, 22)
knitr::kable(t_n_signals_per_chromosome)
```

Now we can get all the associations on 
chromosome `r chromosome_with_most_signals``:

```{r}
t_il_17ra_on_22 <- dplyr::filter(
  t_il_17ra, Chr == chromosome_with_most_signals
)
testthat::expect_equal(nrow(t_il_17ra_on_22), 7)
knitr::kable(t_il_17ra_on_22)
```

Let's see where these are:

```{r fig.width=7, fig.height=1}
ggplot2::ggplot(t_il_17ra_on_22, ggplot2::aes(x = Position., y = 0)) + # nolint period at the end of 'Position', as in original data
  ggplot2::geom_jitter(height = 0.3) + 
  ggplot2::scale_y_continuous(limits = c(-1, 1.0), name = "", labels = NULL)
```

As can be seen, 6 out of 7 SNPs are close to one another,
we'll zoom in now.

First, let's get the primary SNP:

```{r}
t_primary_snp <- t_il_17ra_on_22[ 
  which(t_il_17ra_on_22$X.Primary.Conditional.hit == " Primary"), # nolint extra space in 'Primary', as in original data
]
primary_snp_position <- t_primary_snp$Position.
knitr::kable(t_primary_snp)
```

The primary SNP is located at position `r primary_snp_position`.

Let's create a distance table:

```{r}
distances <- abs(t_il_17ra_on_22$Position. - primary_snp_position)
t_n_in_window <- tibble::tibble(
  n_1_kb    = sum(distances < 1000),
  n_10_kb   = sum(distances < 10000),
  n_100_kb  = sum(distances < 100000),
  n_1000_kb = sum(distances < 1000000)
)
testthat::expect_equal(t_n_in_window$n_1_kb, 2)
testthat::expect_equal(t_n_in_window$n_10_kb, 5)
testthat::expect_equal(t_n_in_window$n_100_kb, 5)
testthat::expect_equal(t_n_in_window$n_1000_kb, 6)
knitr::kable(t_n_in_window)
```

So, there are `r t_n_in_window$n_1_kb` SNPs within a window
of 1 kilobase, which includes the primary SNP.
Subsequently, there are `r t_n_in_window$n_10_kb`,
`r t_n_in_window$n_100_kb` and `r t_n_in_window$n_1000_kb`
SNPs in the 10k, 100k and 1M window respectively.

Note that the 7th SNP is too far away. We'll plot without it:

```{r  fig.width=7, fig.height=1}
ggplot2::ggplot(
  dplyr::filter(
    t_il_17ra_on_22, 
    X.Primary.Conditional.hit != "Conditional 6"
  ),
  ggplot2::aes(x = Position., y = 0)) + # nolint Period at the end of 'Position', as in original data
  ggplot2::geom_jitter(height = 0.3) + 
  ggplot2::scale_y_continuous(limits = c(-1, 1.0), name = "", labels = NULL)
```

The first SNP is now furthest aways, so let's zoom in again:

```{r fig.width=7, fig.height=1}
ggplot2::ggplot(
  dplyr::filter(
    t_il_17ra_on_22, 
    X.Primary.Conditional.hit != "Conditional 6" & X.Primary.Conditional.hit != "Conditional 5"
  ),
  ggplot2::aes(x = Position., y = 0)) + # nolint Period at the end of 'Position', as in original data
  ggplot2::geom_jitter(height = 0.3) + 
  ggplot2::scale_y_continuous(limits = c(-1, 1.0), name = "", labels = NULL)
```
